---
title: "Home Depot Product Search Relevance: Exploratory Analysis"
output: html_document
---

# Defining the necessary directories/importing of packages
```{r}
# Define directories
rawDataFolder <- "../Data/Raw Data/"
processDataFolder <- "../Data/Processed Data/"

# Define raw data
fileName_attributes <- paste0(rawDataFolder, "attributes.csv")
fileName_train <- paste0(rawDataFolder, "train.csv")
fileName_product_descriptions <- paste0(rawDataFolder, "product_descriptions.csv")
```

# Exploring attributes dataset
Take note that the attributes dataset contains ~2 million rows.
- The structure of the data is one of a flat table
- Columns within this data
  - product_uid
  - name
  - value

```{r, cache=TRUE}
attributes <- read.csv(fileName_attributes)

# How many products am I dealing with?
noOfProducts <- unique(attributes$product_uid)
length(noOfProducts) # 86264

# How many types of 'attributes' in this dataset?
attributeType <- unique(attributes$name)
length(attributeType) # 5411 types of product attributes
sort(attributeType) # To view the type of attributes used
```

# Exploring train dataset
- Columns within this dataset
  - id
  - product_uid
  - product_title
  - search_term
  - relevance

```{r, cache=TRUE}
train <- read.csv(fileName_train)

# How many of types of products within this dataset
noOfProducts <- unique(train$product_uid)

```

# Exploring product descriptions
- Columns within this dataset
  - product_uid
  - product_description

```{r, cache=TRUE}
product_descriptions <- read.csv(fileName_product_descriptions)
```

