---
title: "Data Slicing"
output: html_document
---

# Data Slicing
The art of spliting a dataset to its respective training and testing sets. How does one do it? Why do a certain way of data slicing and not the others e.g. Why use bootstraping instead of using n-fold cross validation. Is there any benefit of using either methods?

## A brief summary of how-to-do modelling projects
As with every modelling project, there are several parts before building a model;
* Data Slicing (Spliting the data with at least a 70:30 ratio). 70% of the data is to be used for training purposes while 30% would be used for testing purposes. This would usually be the usual configuration but out there, some possible configurations could be 70:20:10 (training:testing:evaluation).

## Doing a normal split of 70:30 ratio
This way of data slicing involves separating the data via sampling to 2 datasets. 70% of data records fall into the training dataset and 30% of data records fall into the testing dataset.

```{r}
library(caret)
library(kernlab)
# Spam dataset has 4601 rows and 58 columns including the column that is to be predicted
data(spam)
dim(spam)
# Partition is the split percentage
# List is to choose output of whether it being a list or matrix
inTrain <- createDataPartition(y = spam$type, p = 0.7, list = FALSE)
training <- spam[inTrain, ]
testing <- spam[-inTrain,  ]
```

## n-fold cross validation
Cross validation

```{r}
# K-fold data splitting - n-fold cross validation
library(caret)
library(kernlab)
set.seed(1000)
folds <- createFolds(y = spam$type, k = 10, list = TRUE, returnTrain = TRUE)
sapply(folds, length)
folds <- createFolds(y = spam$type, k = 10, list = TRUE, returnTrain = FALSE)
sapply(folds, length)
```

## Bootstrapping

```{r}
# Bootstrapping - sampling with replacement
library(caret)
library(kernlab)
set.seed(1000)
folds <- createResample(y = spam$type, times = 20, list = TRUE)
sapply(folds, length)
```

## Time slicing

```{r}
# Time slices - data that is dependent on order
set.seed(1000)
timeVector <- 1:1000
folds <- createTimeSlices(y = timeVector, initialWindow = 20, horizon = 10)
names(folds)
```

