---
title: "Regression"
output: html_document
---

This document contains sample code on using the caret package to run regression models on data.

## A brief summary of how-to-do modelling projects
As with every modelling project, there are several parts before building a model;
* Data Slicing (Spliting the data with at least a 70:30 ratio). 70% of the data is to be used for training purposes while 30% would be used for testing purposes. This would usually be the usual configuration but out there, some possible configurations could be 70:20:10 (training:testing:evaluation).
* Data exploration. Exploring the data and creating new predictors if necessary (Possible to extract data from already available data and put it into the data)
* Data training. Training the data based on the new preprocessed dataset to predict the 'y' variable
* Model evaluation. Evaluate the accuracy of the model. If you are using logistic regression etc, it may be possible to vizualize this via a ROC curve. A ROC curve > 80% is considered good while anything below 50% is considered worst than random. At this stage, use the testing data available to test the model created from the training data.
* Model deployment. Deploy model to production

## Contents of this document
This document would contain from start to end the steps to do a modelling project and will include the easy and difficult sample codes for easier reference in the future.

List of datasets and a description of its predictors and its aim

Name of dataset | Description of predictors | Aim of modelling this dataset
--- |--- |---
spam dataset in kernlab package | test | Try to predict whether an email is spam or ham based on various predictors available in the dataset

## SPAM Example

```{r, cache = TRUE}
library(caret)
library(kernlab)
data(spam)
# Partition is the split percentage
# List is to choose output of whether it being a list or matrix
inTrain <- createDataPartition(y = spam$type, p = 0.7, list = FALSE)
training <- spam[inTrain, ]
testing <- spam[-inTrain,  ]
dim(training)
# Set seed
# In all computers, random is not really random but dependent on a set of numbers and parameters. One of the parameter is seed. In order to ensure that our model can be replicated by other people, we should set a seed which would allow their results to be comparable to ours
set.seed(1000)
modelFit <- train(type ~ ., data = training, method = "glm") # Generalized linear models
modelFit$finalModel
predictions <- predict(modelFit, newdata = testing)
predictions
confusionMatrix(predictions, testing$type)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
