---
title: "Time Series Analysis"
output: html_document
---

Time series analysis can be done with native R objects such as vector, matrix and data.frame objects. However, in some cases, the information in those native object may contain all the functionality required to do a full analysis on a time series dataset

```{r}
#install.packages("zoo")
library(zoo)

# Getting the data to play around with zoo package
# install.packages("quantmod")
library(quantmod)
library(dplyr)
chartSeries(AAPL)
timeData <- data.frame(AAPL)
timeData <- mutate(timeData, Date = row.names(timeData))
timeData <- select(timeData, Date, AAPL.Close)

# Passing the data to zoo
aapl <- read.zoo(timeData)
```

Understanding what is within the aapl zoo dataset

```{r}
plot(aapl, main = "Apple Closing Prices on NASDAQ", ylab = "Prices (USD)", xlab = "Date")

# Get the data point where aapl was the maximum
aapl[which.max(aapl)]

# Get the data point where aapl was the minimum
aapl[which.min(aapl)]

ret_simple <- diff(aapl) / lag(aapl, k = -1) * 100
ret_count <- diff(log(aapl)) * 100

# coredata strips of all "time" attributes within the xts dataset
summary(coredata(ret_simple))

quantile(ret_simple, probs = 0.01)

# Histogram of returns
hist(ret_simple, breaks = 100, main = "Histogram of Simple Returns", xlab = "%")
```

Restricting the analysis to a shorter time frame

```{r}
aapl_2013 <- window(aapl, start = "2013-01-01", end = "2013-12-31")
aapl_2013[which.max(aapl_2013)]
```

Using the forecast package for modelling time series data

```{r}
```
