---
title: "Understanding the Data Structure"
output: html_document
---

This documents contains possible commands that can be used to further understand the data before starting modelling work etc.

Quick commands to get basic information regarding data.
Some things to probably to try to understand would be:
* Spread of the data
* Whether data has a lot outlier
* How does the each attribute in the data relate to each other and how they relate to the dependent variable

```{r}
str(mtcars)
summary(mtcars)
range(mtcars$mpg)
IQR(mtcars$mpg)
quantile(mtcars$mpg)
quantile(mtcars$hp, probs = c(0.01, 0.99))
```

Plots made at this stage are mainly exploratory in nature. The faster to plot, the faster to understand the data

Some possible plots
```{r}
# Boxplot
boxplot(mtcars$disp, main = "Boxplot of displacement of mtcars", ylab = "cc")

# Histogram plot
hist(mtcars$disp, main = "Histogram of displacement of mtcars", xlab = "cc")
```

## Understanding another dataset: Wage dataset from ISLR

```{r}
# Load libraries
library(ISLR)
library(ggplot2)
library(caret)
library(Hmisc)

# You can immediately identify a few things regarding this data
# Gender is all male from middle atlantic
data(Wage)
summary(Wage)

# Combine this data exploration with concepts from training a model - need to do data slicing
set.seed(1000)
inTrain <- createDataPartition(y = Wage$wage, p = 0.7, list = FALSE)
training <- Wage[inTrain, ]
testing <- Wage[-inTrain, ]
dim(training)
dim(testing)

# Types of plots
# For classification problems - box, strip, density, pairs, ellipsse
# For regression - pairs, scatter
featurePlot(x = training[, c("age", "education", "jobclass")], y = training$wage, plot = "pairs")

# Can use ggplot2 in this case
qplot(age, wage, data = training)
qplot(age, wage, color = jobclass, data = training)

qq <- qplot(age, wage, color = education, data = training)
qq + geom_smooth(method = "lm", formula = y ~ x)

# Let's say you go through numerical datasets
# And there is a need to categorize them, can use the Hmisc package - broken via quantile groups
cutWage <- cut2(training$wage, g = 3)
table(cutWage)

# Boxplot with cut 2
p1 <- qplot(cutWage, age, data = training, fill = cutWage, geom = c("boxplot","jitter"))
p1

# Table data
t1 <- table(cutWage, training$jobclass)
t1
prop.table(t1, 1) # Get proportion of the group - 1 means row, 1 mean group

# Density plots
qplot(wage, color = education, data = training, geom = "density")
```

## Things to look out for during data exploration
* Imbalance in outcomes/predictors (skewed data)
* Outliers
* Groups of points not explained by a predictor
* Skewed variables